---
title: "hw10"
author: "Yonghyun Kwon"
date: "4/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library("lmerTest")
library(NLRoot)
```

# Problem 2

## (a)
```{r}
dat <- read.csv("https://dnett.github.io/S510/hw10GenotypeYield.txt", sep = "\t")
dat$genotype <- factor(dat$genotype)

#(a)
(model1 <- lm(yield ~ 0 + genotype, data = dat))
BLUE <- model1$coefficients
```

## (b)
```{r}
(model2 <- lmer(yield ~ 0 + 1|genotype, data = dat))
#sigma_g^2 = 2.687^2 = 7.22
#sigma_e^2 = 9.669^2 = 93.49
```
$\sigma_g^2$ = 2.687^2 = 7.22 while $\sigma_e^2$ = 9.669^2 = 93.49


## (c)
```{r}
BLUP <- unlist(ranef(model2)) + fixef(model2)
names(BLUP) <- names(BLUE)
BLUP
```

## (d)
```{r}
plot(BLUE ~ BLUP)
abline(0, 1)
```

We can see that the scatters has higher slope than 1. This is because BLUEs have higher variability than BLUPs.

## (e)
```{r}
sort(BLUE, decreasing = TRUE)[1:5]
```

## (f)
```{r}
sort(BLUP, decreasing = TRUE)[1:5]
```

# Problem 3
```{r}
y = c(14, 9, 10, 5, 18, 9, 9, 17, 10, 17,18,13, 17, 16)
G = rep(c(1,2),each = 7)
G = as.factor(G)

model1 = glm(y ~ 0 + G, family = "poisson")
s1 = summary(model1)

model2 = glm(y ~ 1, family = "poisson")
s2 = summary(model2)
```

##(a)
```{r}
AIC(model1)
```

##(b)
```{r}
BIC(model1)
```

##(c)
```{r}
AIC(model2)
```

##(d)
```{r}
BIC(model2)
```

##(e)
According to AIC, the first model is preferred.

##(f)
According to BIC, the first model is preferred.

##(g)
```{r}
- 2 * logLik(model2) + 2 * logLik(model1)
```

##(h)
```{r}
aov <- anova(model1, model2, test = "LRT")
aov$`Pr(>Chi)`[2]
```

##(i)
```{r}
V <- s1$cov.unscaled
thetahat <- s1$coefficients[,1]
(thetahat[2] - thetahat[1]) / sqrt(V[1,1] + V[2,2])
```

##(j)
```{r}
aov <- anova(model1, model2, test = "Rao")
aov$`Pr(>Chi)`[2]
```

##(g)
While the BLUPs are the weighted average of all the sample means and so BLUPs have smaller variablities in this case, the BLUE's are simply the average of i-th genotype.

# Problem 5
## (a)
```{r}
phat = 0.17
n = 100
margin = 1.96 * sqrt(phat * (1-phat) / n)
phat - margin; phat + margin 
```

## (b)
Find the solution of non-linear equation using bisection method in "NLRoot" package.
```{r}
ftn1 <- function(x){
  17 * log(x) + 83 * log(1 - x)
}

lphat = ftn1(phat)
lik2 = lphat - 1/2*qchisq(1- 0.05, 1)

ftn2 <- function(x){
  17 * log(x) + 83 * log(1 - x) - lik2
}

BFfzero(ftn2, 0, phat)
#BFfzero(ftn2, phat, 1) #0.2515018
```

We can see that the confidence interval is (0.1051833, 0.251507)

# Problem 6
We can construct a confidence interval using the following asymptotic distribution
$$
\frac{\hat{p}_1- \hat{p}_2 - (p_1 - p_2)}{\sqrt{Var(\hat{p}_1 - \hat{p}_1)}} \rightarrow N(0,1)
$$

```{r}
p1 = 172 / 350
p2 = 137 / 350
n = 350
margin = 1.96 * sqrt(p1*(1-p1) / n + p2*(1-p2) / n)
p1 - p2 - margin
p1 - p2 + margin
```

So the 95 % confidence interval for $p_1 - p_2$ is(0.02680293, 0.1731971)