---
title: "Hw 7 Problem1"
author: "Kwon Yonghyun"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(emmeans)
```

# Data loading

```{r}
data(pigs)
pigs2 <- pigs
pigs2$percent <- as.factor(pigs2$percent)
pigs2$conc <- log(pigs2$conc)
pigs$conc <- log(pigs$conc)
```

# (a)

To get Type I SS, we can use *lm* function and *anova* function in *R*.

```{r}
model1 <- lm(conc ~ source * percent, data = pigs2)
(anova1 <- anova(model1))
```

# (b)

To obtain Type II SS, we can fit another model and use that Sum of Squares to obtain Sum of Squares for $source | percent, 1$.

```{r}
model2 <- lm(conc ~ percent * source, data = pigs2)
anova2 <- anova(model2)

anova3 <- anova1
anova3[1,] <- anova2[2,]
anova3
```

# (c)

To obtain Type III SS, we can use *joint_tests* and *emmeans* functions.

```{r}
anova4 <- joint_tests(emmeans(model1, specs = c("source", "percent")))
MSE = anova1["Residuals", "Mean Sq"]
MSR = anova4$F.ratio * MSE
SumSq = anova4$df1 * MSR
anova4 <- cbind(Df = anova4$df1, SumSq = SumSq, MeanSq = MSR, Fvalue = anova4$F.ratio, pval = anova4$p.value)
anova4 <- rbind(anova4, unlist(anova3[4,]))
anova4
```

Note that *joint_test* function does not give Sum of Squares. Hence, we have to manually compute Sum of Squares and Mean Sum of Squares in R code.

# (d)

Ls means can be computed using *tapply* function.

```{r}
tapply(pigs2$conc, pigs2$source, mean)
tapply(pigs2$conc, pigs2$percent, mean)
```

# (e)

```{r}
model3 <- lm(conc ~ source * percent, data = pigs)
anova(model1, model3)
```

p-value greater than 0.05 implies that there is no strong evidence to reject the null hypothesis. Therefore, we can say that such linear model fit adequately relatvie to the cell-means model.

# (f)

Estimated linear relationship for each source can be represented as

\begin{align*}
fish &: \mu_R + x \beta_R \\
soy &: \mu_R + \alpha_{Rs} + x( \beta_R + \gamma_{Rs}) \\
skim &: \mu_R + \alpha_{Rk} + x( \beta_R + \gamma_{Rk})
\end{align*}

Hence, the estimated relationship is
```{r}
coef <- model3$coefficients
sprintf("fish: y = %2.2f + x * %2.2f", coef[1], coef[4])
sprintf("soy: y = %2.2f + x * %2.2f", coef[1] + coef[2], coef[4] + coef[5])
sprintf("skim: y = %2.2f + x * %2.2f", coef[1] + coef[3], coef[4] + coef[6])
```


